---
title: "data_mb"
author: "Martine Besse"
date: "2023-05-12"
output: html_document
---
# Environment
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("readr") 
library("sf")
library("ggplot2")
library("dplyr")
library("lubridate") #for timezones
library("raster")
library("leaflet")
```

# Posmo data exploration

## Import, clean and add useful attributes
### Clean posmo
```{r}
#import and convert to sf object
posmo <- read_csv("data/posmo_2023-06-14_2.csv") %>% st_as_sf(coords = c("lon_x", "lat_y"), crs=4326, remove = FALSE) |> st_transform(2056)

#clean and add columns
posmo$source <- "posmo"

posmo <- posmo %>% 
       rename("tmode_posmo" = "transport_mode")

posmo$tmode_manual <- NA

to_remove <- c("place_name", "lon_x", "lat_y", "user_id")
posmo <- posmo[ , !(names(posmo) %in% to_remove)]

#add East and North columns
coords <- posmo |> st_coordinates()
posmo <- posmo |> 
  mutate(E =  coords[,1], N = coords[,2])

#update timezone
posmo <- posmo %>%
  mutate(
    datetime = datetime %>% with_tz(tzone = "Europe/Zurich")
  )

```

### Add new variables
#### Compute new values
```{r}
#compute new values
#timelag
posmo <- mutate(posmo, timelag = as.numeric(difftime(lead(datetime), datetime)))

#steplength
posmo <- posmo %>% mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2))

#speed
posmo <- posmo %>% mutate(speed_ms = steplength_m/timelag)
```


<!-- #### Add heights -->
<!-- The height comes from the Digital height model DHM25 provided by swisstopo. The raster has been preprocessed on QGIS to be converted into the epsg 2056 (previously 21781) to save calculation time.   -->
<!-- The resolution is 25m. To have more precise information, we could look at the swiss3DAlti if needed.   -->
<!-- ```{r} -->
<!-- #Extract values from rasters: -->
<!-- #https://gisday.wordpress.com/2014/03/24/extract-raster-values-from-points-using-r/comment-page-1/ -->
<!-- dhm25 <- raster("./data/dhm25_raster/dhm25_2056.tif") -->
<!-- rasValue <- extract(dhm25, posmo) -->
<!-- posmo$H <- rasValue -->
<!-- ``` -->

<!-- ```{r} -->
<!-- plot(dhm25) -->
<!-- plot(posmo) -->
<!-- ``` -->

<!-- Simplification: we did not consider height of soil compared to height of person or public transport.   -->
<!-- As it is a relative variable to calculate slope, it does not matter.   -->

<!-- ```{r} -->
<!-- # Add slope -->
<!-- posmo <- posmo %>% mutate(slope = (lead(H) - H)/steplength_m * 100) -->
<!-- ``` -->



<!-- ### Reorder columns -->
<!-- ```{r} -->
<!-- col_order <- c("source", "datetime", "weekday", "tmode_posmo", "tmode_manual" ,"E", "N", "H", "timelag", "steplength_m", "speed_ms", "slope", "geometry") -->
<!-- posmo <- posmo[, col_order] -->
<!-- posmo -->
<!-- ``` -->


## Overview
### Overview map
```{r}
ggplot() +
  geom_path(data = posmo, aes(x = E, y = N), colour = 'red') +
  ggtitle("My trajectories from starting to record data 30.03.2023 to now") +
  coord_equal()
```

```{r}
#plot only points
ggplot(posmo) +
  geom_point(aes(E, N, colour = datetime)) +
  coord_equal() #so equally space North and East
```


```{r}
posmo_wgs <- posmo %>% st_transform(crs = 4326)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(data = posmo_wgs,
                   #opacity = 0.3, 
                   radius = 0.2,
                   popup = posmo_wgs$datetime,
                   color = 'blue') %>%
  addLegend(position = 'topright', 
            colors = c('blue'), 
            labels = c('points'),
            title = 'Downloaded points from posmo')

# Plot the leaflet object m
m
```

### Overview time
```{r}
#extract Hour of the day
posmo_hour <- posmo |> 
  mutate(
    H = as.numeric(format(as.POSIXct(posmo$datetime), format = "%H"))
  )

ggplot(posmo_hour, aes(H)) +
  geom_histogram(binwidth = .6)
```


### Select one single date
Select one single date:  
```{r}
posmo_filter <- posmo |>
    filter(as.Date(datetime) == "2023-05-4")
posmo_filter
```

```{r}
#plot trajectories
ggplot(posmo_filter, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


## Clean data
(remove steps when no recorded, static etc.)  
remove statics: Exercice 3  
Add segment IDs: exercice 3 - task 4 
(similarity measure ex3)
assigning 

## Visualise per zone
Boundaries:  
Zurich: lower left: 2674900.0, 1244000.0  
        upper right: 2689600.0, 1257300  

Waadt: 	lower left: 2480000, 1127600  
        upper right: 2575200, 1195100  
```{r}
# add regions
#Zurich
zh_ll <- c(2674900, 1244000)
zh_ur <- c(2692100, 1260300)

#Waadt
wt_ll <- c(2480000, 1127600)
wt_ur <- c(2575200, 1195100)

#wallis
ws_ll <- c(2538000, 1076875)
ws_ur <- c(2661500, 1153125)


posmo$region <- NA
posmo <- posmo |> 
  mutate(
    region = case_when(
      E > zh_ll[1] & E < zh_ur[1] & N > zh_ll[2] & N < zh_ur[2] ~ 'Zurich', 
      E > wt_ll[1] & E < wt_ur[1] & N > wt_ll[2] & N < wt_ur[2] ~ 'Waadt',
      E > ws_ll[1] & E < ws_ur[1] & N > ws_ll[2] & N < ws_ur[2] ~ 'Wallis'
      )
    )

posmo_zh <- posmo %>% filter(region == "Zurich")
posmo_waadt <- posmo %>% filter(region == "Waadt")
posmo_wallis <- posmo %>% filter(region == "Wallis")

ggplot(posmo_wallis, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


```{r}
posmo_wgs <- posmo_zh %>% st_transform(crs = 4326)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(data = posmo_wgs,
                   #opacity = 0.3, 
                   radius = 0.2,
                   popup = posmo_wgs$datetime,
                   color = 'blue') %>%
  addLegend(position = 'topright', 
            colors = c('blue'), 
            labels = c('points'),
            title = 'Downloaded points from posmo')

# Plot the leaflet object m
m
```


# Labelize data
tendency per labelized data (mean speed, mean distance, etc.)

Labels:  
car, train, bus, tram, cycling, kick-scooter, walk, ski, ski_lift, boat
To be sure to cut out unwanted parts, add a margin to the sample.  

## Boats
```{r}
#BOAT (14.05)
#2023-05-14 13:30:00
#2023-05-14 14:42:00

condition_boats <- posmo$datetime > as.POSIXct("2023-05-14 13:55:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-14 14:21:00", tz = "Europe/Zurich")

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_boats ~ 'boat'))

# Visualisation of all labelized tmode datasets
posmo_boats <- posmo %>% filter(tmode_manual == "boat")

#plot trajectories
ggplot(posmo_boats, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Ski
-15.04, 
-13.04, 
12.04
### Cable car
```{r}
condition_cable_cars <- 
  posmo$datetime > as.POSIXct("2023-04-12 09:12:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:18:30", tz = "Europe/Zurich")

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_cable_cars ~ 'cable_car'))

# Visualisation of all labelized tmode datasets
posmo_cable_cars <- posmo %>% filter(tmode_manual == "cable_car")

#plot trajectories
ggplot(posmo_cable_cars, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

### Ski lift
```{r}
condition_ski_lifts <- 
  posmo$datetime > as.POSIXct("2023-04-12 09:23:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:41:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:51:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 9:59:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:15:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:24:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:51:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:57:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:08:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:21:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:29:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:46:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:51:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 13:49:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 13:54:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 13:57:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 14:05:00", tz = "Europe/Zurich")

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_ski_lifts ~ 'ski_lift'))

# Visualisation of all labelized tmode datasets
posmo_ski_lifts <- posmo %>% filter(tmode_manual == "ski_lift")

#plot trajectories
ggplot(posmo_ski_lifts, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


### T-bar
```{r}
condition_t_bars <- 
  posmo$datetime > as.POSIXct("2023-04-12 10:28:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:33:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 14:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 14:18:00", tz = "Europe/Zurich") 

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_t_bars ~ 't_bar'))

# Visualisation of all labelized tmode datasets
posmo_t_bars <- posmo %>% filter(tmode_manual == "t_bar")

#plot trajectories
ggplot(posmo_t_bars, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

### Ski slopes
```{r}
condition_skis <- 
  posmo$datetime > as.POSIXct("2023-04-12 09:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:33:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:41:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 9:51:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:59:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:15:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:24:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:28:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:51:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:57:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:00:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:08:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:21:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:29:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:46:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:51:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 13:49:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 13:54:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 13:57:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 14:05:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 14:10:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 14:18:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 14:48:00", tz = "Europe/Zurich") 


posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_skis ~ 'ski'))

# Visualisation of all labelized tmode datasets
posmo_skis <- posmo %>% filter(tmode_manual == "ski")

#plot trajectories
ggplot(posmo_skis, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Kick-scooter
- 27.04 -> no data
- 3.05 -> vélo
4.05
- 24.05 (le soir 19h)
```{r}
#select bikes segments
condition_scooters <- 
  posmo$datetime > as.POSIXct("2023-05-04 07:36:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-04 07:42:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-24 22:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-24 22:10:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-24 19:48:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-24 20:00:00", tz = "Europe/Zurich")
  
  
#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_scooters ~ 'kick-scooter'))

# Visualisation of all labelized tmode datasets
posmo_scooters <- posmo %>% filter(tmode_manual == "kick-scooter") #%>% filter(region == "Zurich") #just for visualization, bc only one trajectory in Lausanne

#plot trajectories
ggplot(posmo_scooters, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Cycling
02.05
-03.05,
15.05,
-17.05,
-23.05,
-24.05,
-25.05,
-31.05,
-1.06

```{r}
#select bikes segments
condition_bikes <- 
  posmo$datetime > as.POSIXct("2023-05-02 08:31:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-02 08:36:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-02 11:49:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-02 12:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-02 17:44:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-02 18:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-15 15:57:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-15 16:32:00", tz = "Europe/Zurich") 

#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_bikes ~ 'bike'))

# Visualisation of all labelized tmode datasets
posmo_bikes <- posmo %>% filter(tmode_manual == "bike") %>% filter(region == "Zurich") #just for visualization, bc only one trajectory in Lausanne

#plot trajectories
ggplot(posmo_bikes, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


## Railways
TO TAKE: -

-02.06
26.05
09.05
-05.05
28.04
-04.04
03.04
-31.03 -> to take (but map matched)
30.03

```{r}
#select train segments
condition_trains <- 
  posmo$datetime > as.POSIXct("2023-03-30 17:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-03-30 17:52:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-05-26 18:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-26 19:15:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-05-09 8:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-09 10:55:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-04-28 12:20:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-28 12:50:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-06-09 19:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 21:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-03 22:15:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-04 00:22:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-09 12:23:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 12:48:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-11 02:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-11 02:45:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-12 07:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-12 07:53:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-06-14 17:25:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-14 17:42:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 09:05:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 11:00:00", tz = "Europe/Zurich")
  
#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_trains ~ 'train'))

# Visualisation of all labelized rails datasets
posmo_rails <- posmo %>% filter(tmode_manual == "train")

#plot trajectories
ggplot(posmo_rails, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```



## Trams
- 22.05
- 13.05
- 25.04

```{r}
#select trams segments
condition_trams <- 
  posmo$datetime > as.POSIXct("2023-04-16 15:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-16 15:37:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-06-09 17:17:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 17:22:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-09 13:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 13:54:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 11:09:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 11:23:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-05 18:49:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-05 19:03:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-05 23:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-05 23:40:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 15:50:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 15:57:00", tz = "Europe/Zurich")

  
#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_trams ~ 'tram'))

# Visualisation of all labelized tmode datasets
posmo_trams <- posmo %>% filter(tmode_manual == "tram") %>% filter(region == "Zurich") #Waadt #just for visualization, 

#plot trajectories
ggplot(posmo_trams, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


## Bus
- 03.05
- 02.05
25.04
```{r}
#select bus segments
condition_bus <- 
  posmo$datetime > as.POSIXct("2023-06-09 17:25:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 17:31:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-24 19:32:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-24 19:49:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-24 22:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-24 22:22:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-07 07:53:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-07 08:12:00", tz = "Europe/Zurich") | posmo$datetime > as.POSIXct("2023-06-07 08:35:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-07 08:45:00", tz = "Europe/Zurich") | posmo$datetime > as.POSIXct("2023-06-07 19:23:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-07 19:37:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 20:39:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 20:44:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 21:03:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 21:17:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 08:15:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 08:25:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 08:28:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 08:35:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-05 18:37:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-05 18:46:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 11:36:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 11:57:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 16:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 16:39:00", tz = "Europe/Zurich")

#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_bus ~ 'bus'))

# Visualisation of all labelized tmode datasets
posmo_bus <- posmo %>% filter(tmode_manual == "bus") %>% filter(region == "Zurich") #Waadt #just for visualization, 

#plot trajectories
ggplot(posmo_bus, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```



## Walking
29.05
22.05
-21.05
20.05
13.05
16.04

```{r}
#select walking segments
condition_walks <- 
  posmo$datetime > as.POSIXct("2023-04-16 15:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-16 15:37:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-04-16 15:44:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-16 15:56:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-13 13:29:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-13 15:01:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-13 16:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-13 22:01:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-13 22:04:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-13 22:29:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-20 11:54:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-20 22:35:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-22 11:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-22 11:42:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-29 16:18:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-29 16:46:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-09 11:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 12:20:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-11 02:46:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-11 03:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-12 07:53:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-12 08:07:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 08:11:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 08:15:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-05 23:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-05 23:57:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:12:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 18:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 18:50:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 11:57:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 12:05:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 15:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 15:50:00", tz = "Europe/Zurich") 


#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_walks ~ 'walk'))

# Visualisation of all labelized tmode datasets
posmo_walks <- posmo %>% filter(tmode_manual == "walk") %>% filter(region == "Zurich") #Waadt #just for visualization, 

#plot trajectories
ggplot(posmo_walks, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Car
- 30.04
- 25.05
24.04
- 22.04
- 17.04
- 15.04

```{r}
#select car segments
condition_cars <- 
  posmo$datetime > as.POSIXct("2023-06-11 15:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-11 17:00:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-04-24 14:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-24 17:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-10 13:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-10 17:33:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-09 21:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 22:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-14 07:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-14 08:00:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-14 17:44:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-14 18:10:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-13 18:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-13 19:15:00", tz = "Europe/Zurich") 
  
#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_cars ~ 'car'))

# Visualisation of all labelized tmode datasets
posmo_cars <- posmo %>% filter(tmode_manual == "car") #%>% filter(region == "Zurich") #Waadt #just for visualization, 

#plot trajectories
ggplot(posmo_cars, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## test
```{r}
# to see segments

#tram
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-25 15:50:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 15:57:00", tz = "Europe/Zurich"))
posmo_test



posmo_wgs <- posmo_test %>% st_transform(crs = 4326)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(data = posmo_wgs,
                   #opacity = 0.3, 
                   radius = 0.2,
                   popup = posmo_wgs$datetime,
                   color = 'blue') %>%
  addLegend(position = 'topright', 
            colors = c('blue'), 
            labels = c('points'),
            title = 'Downloaded points from posmo')

# Plot the leaflet object m
m

# #plot trajectories
# ggplot(posmo_test, aes(E,N, color = datetime)) +
#   geom_point() +
#   geom_path() +
#   coord_equal() #so equally space North and East
```


```{r}
#final labelizing
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_walks ~ 'walk',
                                  condition_boats ~ 'boat',
                                  condition_trains ~ 'train',
                                  condition_bikes ~ 'bike',
                                  condition_scooters ~ 'kick-scooter',
                                  condition_cars ~ 'car',
                                  condition_bus ~ 'bus',
                                  condition_trams ~ 'tram',
                                  condition_t_bars ~ 't_bar',
                                  condition_cable_cars ~ 'cable_car',
                                  condition_ski_lifts ~ 'ski_lift',
                                  condition_skis ~ 'ski'
                                  ))


posmo %>% group_by(tmode_manual) %>% count()


posmo_filter <- posmo %>% filter(tmode_manual == "ski") #%>% filter(!is.na(tmode_manual)) #%>% filter(region == "Zurich") 

posmo_wgs <- posmo_filter %>% st_transform(crs = 4326)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(data = posmo_wgs,
                   #opacity = 0.3, 
                   radius = 0.2,
                   popup = posmo_wgs$datetime,
                   color = 'blue') %>%
  addLegend(position = 'topright', 
            colors = c('blue'), 
            labels = c('points'),
            title = 'Downloaded points from posmo')

# Plot the leaflet object m
m


```


# Qualify our labelized dataset 
Deriving speed: Ex2 - task5 (rolling window)  

