---
title: "data_mb"
author: "Martine Besse"
date: "2023-05-12"
output: html_document
---
# Environment
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("readr") 
library("sf")
library("ggplot2")
library("dplyr")
library("lubridate") #for timezones
library("raster")
library("leaflet")
```

# Posmo data exploration

## Import, clean and add useful attributes
### Clean posmo
```{r}
#import and convert to sf object
posmo <- read_csv("data/posmo_2023-06-18.csv") %>% st_as_sf(coords = c("lon_x", "lat_y"), crs=4326, remove = FALSE) |> st_transform(2056)

#clean and add columns
posmo$source <- "posmo"

posmo <- posmo %>% 
       rename("tmode_posmo" = "transport_mode")

posmo$tmode_manual <- NA

to_remove <- c("place_name", "lon_x", "lat_y", "user_id")
posmo <- posmo[ , !(names(posmo) %in% to_remove)]

#add East and North columns
coords <- posmo |> st_coordinates()
posmo <- posmo |> 
  mutate(E =  coords[,1], N = coords[,2])

#update timezone
posmo <- posmo %>%
  mutate(
    datetime = datetime %>% with_tz(tzone = "Europe/Zurich")
  )

```

### Add new variables
#### Compute new values
```{r}
#compute new values
#timelag
posmo <- mutate(posmo, timelag = as.numeric(difftime(lead(datetime), datetime)))

#steplength
posmo <- posmo %>% mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2))

#speed
posmo <- posmo %>% mutate(speed_ms = steplength_m/timelag)
```


<!-- #### Add heights -->
<!-- The height comes from the Digital height model DHM25 provided by swisstopo. The raster has been preprocessed on QGIS to be converted into the epsg 2056 (previously 21781) to save calculation time.   -->
<!-- The resolution is 25m. To have more precise information, we could look at the swiss3DAlti if needed.   -->
<!-- ```{r} -->
<!-- #Extract values from rasters: -->
<!-- #https://gisday.wordpress.com/2014/03/24/extract-raster-values-from-points-using-r/comment-page-1/ -->
<!-- dhm25 <- raster("./data/dhm25_raster/dhm25_2056.tif") -->
<!-- rasValue <- extract(dhm25, posmo) -->
<!-- posmo$H <- rasValue -->
<!-- ``` -->

<!-- ```{r} -->
<!-- plot(dhm25) -->
<!-- plot(posmo) -->
<!-- ``` -->

<!-- Simplification: we did not consider height of soil compared to height of person or public transport.   -->
<!-- As it is a relative variable to calculate slope, it does not matter.   -->

<!-- ```{r} -->
<!-- # Add slope -->
<!-- posmo <- posmo %>% mutate(slope = (lead(H) - H)/steplength_m * 100) -->
<!-- ``` -->



<!-- ### Reorder columns -->
<!-- ```{r} -->
<!-- col_order <- c("source", "datetime", "weekday", "tmode_posmo", "tmode_manual" ,"E", "N", "H", "timelag", "steplength_m", "speed_ms", "slope", "geometry") -->
<!-- posmo <- posmo[, col_order] -->
<!-- posmo -->
<!-- ``` -->


## Overview
### Overview map
```{r}
ggplot() +
  geom_path(data = posmo, aes(x = E, y = N), colour = 'red') +
  ggtitle("My trajectories from starting to record data 30.03.2023 to now") +
  coord_equal()
```

```{r}
#plot only points
ggplot(posmo) +
  geom_point(aes(E, N, colour = datetime)) +
  coord_equal() #so equally space North and East
```


```{r}
posmo_wgs <- posmo %>% st_transform(crs = 4326)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(data = posmo_wgs,
                   #opacity = 0.3, 
                   radius = 0.2,
                   popup = posmo_wgs$datetime,
                   color = 'blue') %>%
  addLegend(position = 'topright', 
            colors = c('blue'), 
            labels = c('points'),
            title = 'Downloaded points from posmo')

# Plot the leaflet object m
m
```

### Overview time
```{r}
#extract Hour of the day
posmo_hour <- posmo |> 
  mutate(
    H = as.numeric(format(as.POSIXct(posmo$datetime), format = "%H"))
  )

ggplot(posmo_hour, aes(H)) +
  geom_histogram(binwidth = .6)
```


### Select one single date
Select one single date:  
```{r}
posmo_filter <- posmo |>
    filter(as.Date(datetime) == "2023-05-4")
posmo_filter
```

```{r}
#plot trajectories
ggplot(posmo_filter, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


## Clean data
(remove steps when no recorded, static etc.)  
remove statics: Exercice 3  
Add segment IDs: exercice 3 - task 4 
(similarity measure ex3)
assigning 

## Visualise per zone
Boundaries:  
Zurich: lower left: 2674900.0, 1244000.0  
        upper right: 2689600.0, 1257300  

Waadt: 	lower left: 2480000, 1127600  
        upper right: 2575200, 1195100  
```{r}
# add regions
#Zurich
zh_ll <- c(2674900, 1244000)
zh_ur <- c(2692100, 1260300)

#Waadt
wt_ll <- c(2480000, 1127600)
wt_ur <- c(2575200, 1195100)

#wallis
ws_ll <- c(2538000, 1076875)
ws_ur <- c(2661500, 1153125)


posmo$region <- NA
posmo <- posmo |> 
  mutate(
    region = case_when(
      E > zh_ll[1] & E < zh_ur[1] & N > zh_ll[2] & N < zh_ur[2] ~ 'Zurich', 
      E > wt_ll[1] & E < wt_ur[1] & N > wt_ll[2] & N < wt_ur[2] ~ 'Waadt',
      E > ws_ll[1] & E < ws_ur[1] & N > ws_ll[2] & N < ws_ur[2] ~ 'Wallis'
      )
    )

posmo_zh <- posmo %>% filter(region == "Zurich")
posmo_waadt <- posmo %>% filter(region == "Waadt")
posmo_wallis <- posmo %>% filter(region == "Wallis")

ggplot(posmo_wallis, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


```{r}
posmo_wgs <- posmo_zh %>% st_transform(crs = 4326)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(data = posmo_wgs,
                   #opacity = 0.3, 
                   radius = 0.2,
                   popup = posmo_wgs$datetime,
                   color = 'blue') %>%
  addLegend(position = 'topright', 
            colors = c('blue'), 
            labels = c('points'),
            title = 'Downloaded points from posmo')

# Plot the leaflet object m
m
```


# Labelize data
tendency per labelized data (mean speed, mean distance, etc.)

Labels:  
car, train, bus, tram, cycling, kick-scooter, walk, ski, ski_lift, boat
To be sure to cut out unwanted parts, add a margin to the sample.  

## Boats
```{r}
#BOAT (14.05)
#2023-05-14 13:30:00
#2023-05-14 14:42:00

condition_boats <- posmo$datetime > as.POSIXct("2023-05-14 13:55:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-14 14:21:00", tz = "Europe/Zurich")

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_boats ~ 'boat'))

# Visualisation of all labelized tmode datasets
posmo_boats <- posmo %>% filter(tmode_manual == "boat")

#plot trajectories
ggplot(posmo_boats, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Plane
```{r}
condition_planes <- posmo$datetime > as.POSIXct("2023-06-17 11:12:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-17 12:22:00", tz = "Europe/Zurich")

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_planes ~ 'plane'))

# Visualisation of all labelized tmode datasets
posmo_planes <- posmo %>% filter(tmode_manual == "plane")

#plot trajectories
ggplot(posmo_planes, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


## Ski
-15.04, 
-13.04, 
12.04
### Cable car
```{r}
condition_cable_cars <- 
  posmo$datetime > as.POSIXct("2023-04-12 09:12:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:18:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:02:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 09:10:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 12:20:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 12:27:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 09:18:10", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 9:24:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 13:11:45", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 13:20:00", tz = "Europe/Zurich")

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_cable_cars ~ 'cable_car'))

# Visualisation of all labelized tmode datasets
posmo_cable_cars <- posmo %>% filter(tmode_manual == "cable_car")

#plot trajectories
ggplot(posmo_cable_cars, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

### Ski lift
```{r}
condition_ski_lifts <- 
  posmo$datetime > as.POSIXct("2023-04-12 09:23:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:41:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:51:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 9:59:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:15:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:24:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:51:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:57:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:08:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:21:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:29:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:46:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:51:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 13:49:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 13:54:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 13:57:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 14:05:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:14:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 09:19:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:22:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 09:30:45", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:36:45", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 09:44:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:58:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 10:08:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 10:19:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 10:28:30", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-04-13 10:40:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 10:51:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 09:28:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 9:33:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 09:36:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 9:43:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 09:48:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 9:56:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 10:06:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 10:14:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 10:23:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 10:28:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 10:30:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 10:38:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 11:06:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 11:14:30", tz = "Europe/Zurich")

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_ski_lifts ~ 'ski_lift'))

# Visualisation of all labelized tmode datasets
posmo_ski_lifts <- posmo %>% filter(tmode_manual == "ski_lift")

#plot trajectories
ggplot(posmo_ski_lifts, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


### T-bar
```{r}
condition_t_bars <- 
  posmo$datetime > as.POSIXct("2023-04-12 10:28:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:33:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 14:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 14:18:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 10:54:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 11:02:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 11:19:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 11:26:45", tz = "Europe/Zurich")

posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_t_bars ~ 't_bar'))

# Visualisation of all labelized tmode datasets
posmo_t_bars <- posmo %>% filter(tmode_manual == "t_bar")

#plot trajectories
ggplot(posmo_t_bars, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

### Ski slopes
```{r}
condition_skis <- 
  posmo$datetime > as.POSIXct("2023-04-12 09:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:33:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:41:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 9:51:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:59:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:15:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:24:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:28:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 10:51:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 10:57:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:00:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:08:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:21:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:29:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 11:46:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 11:51:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 13:49:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 13:54:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 13:57:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 14:05:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 14:10:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 14:18:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 14:48:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:12:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 09:14:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:19:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 09:22:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:30:45", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 09:36:45", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 09:44:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 09:58:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 10:08:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 10:17:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 10:28:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 10:40:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 10:51:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 10:54:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 11:02:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 11:25:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 15:40:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 15:59:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 09:24:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 9:28:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 09:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 9:36:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 09:43:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 9:48:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 09:56:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 10:06:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 10:14:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 10:23:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 10:28:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 10:30:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 10:38:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 11:06:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 11:14:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 11:19:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 11:26:45", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 11:42:30", tz = "Europe/Zurich")


posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_skis ~ 'ski'))

# Visualisation of all labelized tmode datasets
posmo_skis <- posmo %>% filter(tmode_manual == "ski")

#plot trajectories
ggplot(posmo_skis, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Kick-scooter
- 27.04 -> no data
- 3.05 -> vélo
4.05
- 24.05 (le soir 19h)
```{r}
#select kick scooters segments
condition_scooters <- 
  posmo$datetime > as.POSIXct("2023-05-04 07:36:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-04 07:42:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-24 22:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-24 22:10:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-24 19:48:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-24 20:00:00", tz = "Europe/Zurich")
  
  
#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_scooters ~ 'kick-scooter'))

# Visualisation of all labelized tmode datasets
posmo_scooters <- posmo %>% filter(tmode_manual == "kick-scooter") #%>% filter(region == "Zurich") #just for visualization, bc only one trajectory in Lausanne

#plot trajectories
ggplot(posmo_scooters, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Cycling
02.05
-03.05,
15.05,
-17.05,
-23.05,
-24.05,
-25.05,
-31.05,
-1.06

```{r}
#select bikes segments
condition_bikes <- 
  posmo$datetime > as.POSIXct("2023-05-02 08:31:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-02 08:36:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-02 11:49:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-02 12:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-02 17:44:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-02 18:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-15 15:57:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-15 16:32:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-01 07:37:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-01 07:53:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-01 18:38:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-01 18:52:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 07:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 08:00:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 17:32:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 17:50:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 18:02:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 18:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 18:09:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 18:35:00", tz = "Europe/Zurich")

#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_bikes ~ 'bike'))

# Visualisation of all labelized tmode datasets
posmo_bikes <- posmo %>% filter(tmode_manual == "bike") %>% filter(region == "Zurich") #just for visualization, bc only one trajectory in Lausanne

#plot trajectories
ggplot(posmo_bikes, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


## Railways
TO TAKE: -

-02.06
26.05
09.05
-05.05
28.04
-04.04
03.04
-31.03 -> to take (but map matched)
30.03

```{r}
#select train segments
condition_trains <- 
  posmo$datetime > as.POSIXct("2023-05-26 18:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-26 19:15:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-05-09 8:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-09 10:55:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-04-28 12:20:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-28 12:50:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-06-09 19:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 21:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-09 12:23:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 12:48:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-11 02:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-11 02:45:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-12 07:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-12 07:53:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-06-14 17:25:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-14 17:42:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 09:05:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 11:00:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-15 08:06:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-15 10:02:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-16 18:48:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-16 19:11:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-18 12:45:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-18 16:33:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-02 07:04:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-02 07:43:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-02 15:27:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-02 17:53:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-03-31 16:23:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-03-31 16:47:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-03 21:32:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-04 00:22:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-06 17:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-06 17:57:00", tz = "Europe/Zurich") 
  
#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_trains ~ 'train'))

# Visualisation of all labelized rails datasets
posmo_rails <- posmo %>% filter(tmode_manual == "train")

#plot trajectories
ggplot(posmo_rails, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```



## Trams
- 22.05
- 13.05
- 25.04

```{r}
#select trams segments
condition_trams <- 
  posmo$datetime > as.POSIXct("2023-06-09 17:17:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 17:22:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-09 13:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 13:54:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 11:09:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 11:23:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-05 18:49:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-05 19:03:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-05 23:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-05 23:40:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 15:50:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 15:57:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-15 10:07:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-15 10:19:45", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-15 18:36:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-15 18:56:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-04 23:05:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-04 23:16:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-04 16:51:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-04 17:01:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-04 19:29:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-04 19:41:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-05 08:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-05 08:52:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-06 07:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-06 07:50:00", tz = "Europe/Zurich") 

  
#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_trams ~ 'tram'))

# Visualisation of all labelized tmode datasets
posmo_trams <- posmo %>% filter(tmode_manual == "tram") %>% filter(region == "Zurich") #Waadt #just for visualization, 

#plot trajectories
ggplot(posmo_trams, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


## Bus
- 03.05
- 02.05
25.04
```{r}
#select bus segments
condition_bus <- 
  posmo$datetime > as.POSIXct("2023-06-09 17:25:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 17:31:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-24 19:32:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-24 19:49:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-24 22:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-24 22:22:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-07 07:53:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-07 08:12:00", tz = "Europe/Zurich") | posmo$datetime > as.POSIXct("2023-06-07 08:35:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-07 08:45:00", tz = "Europe/Zurich") | posmo$datetime > as.POSIXct("2023-06-07 19:23:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-07 19:37:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 20:39:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 20:44:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 21:03:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 21:17:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 08:15:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 08:25:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 08:28:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 08:35:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-05 18:37:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-05 18:46:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 11:36:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 11:57:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 16:33:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 16:39:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-17 16:07:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-17 16:19:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-18 12:12:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-18 12:16:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-18 12:23:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-18 12:29:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-04 23:17:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-04 23:23:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-02 06:59:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-02 07:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-02 07:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-02 07:50:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-01 22:17:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-01 22:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 21:57:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 22:14:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-03-31 16:13:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-03-31 16:20:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-04 19:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-04 19:59:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-06 07:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-06 07:33:00", tz = "Europe/Zurich")

#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_bus ~ 'bus'))

# Visualisation of all labelized tmode datasets
posmo_bus <- posmo %>% filter(tmode_manual == "bus") %>% filter(region == "Zurich") #Waadt #just for visualization, 

#plot trajectories
ggplot(posmo_bus, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```



## Walking
29.05
22.05
-21.05
20.05
13.05
16.04

```{r}
#select walking segments
condition_walks <- 
  posmo$datetime > as.POSIXct("2023-04-16 15:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-16 15:37:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-04-16 15:44:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-16 15:56:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-13 13:29:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-13 15:01:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-13 16:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-13 22:01:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-13 22:04:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-13 22:29:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-20 11:54:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-20 22:35:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-22 11:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-22 11:42:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-29 16:18:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-29 16:46:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-09 11:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 12:20:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-11 02:46:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-11 03:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-12 07:53:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-12 08:07:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-06 08:11:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-06 08:15:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-05 23:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-05 23:57:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 09:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 09:12:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-12 18:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-12 18:50:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 11:57:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 12:05:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-25 15:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-25 15:50:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 08:43:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 08:46:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-13 18:05:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-13 18:15:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 13:20:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 13:25:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-15 10:19:45", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-15 10:24:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-15 18:56:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-15 19:15:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-16 18:20:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-16 18:45:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-16 21:41:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-16 22:23:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-17 15:59:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-17 16:04:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-17 16:20:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-17 16:28:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-17 22:31:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-17 22:45:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-18 12:16:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-18 12:19:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-18 12:29:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-18 12:33:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-04 17:59:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-04 18:12:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-04 22:50:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-04 23:05:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-02 15:04:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-02 15:23:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 17:50:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 18:02:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 18:04:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 18:09:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 19:47:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 20:00:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-05-31 21:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-05-31 21:55:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-10 18:29:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-10 18:45:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-10 22:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-10 22:18:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-11 11:51:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-11 11:56:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-11 15:28:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-11 15:36:00", tz = "Europe/Zurich")


#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_walks ~ 'walk'))

# Visualisation of all labelized tmode datasets
posmo_walks <- posmo %>% filter(tmode_manual == "walk") %>% filter(region == "Waadt") #Waadt #just for visualization, 

#plot trajectories
ggplot(posmo_walks, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Car
- 30.04
- 25.05
24.04
- 22.04
- 17.04
- 15.04

```{r}
#select car segments
condition_cars <- 
  posmo$datetime > as.POSIXct("2023-06-11 15:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-11 17:00:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-04-24 14:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-24 17:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-10 13:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-10 17:33:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-09 21:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-09 22:30:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-14 07:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-14 08:00:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-14 17:44:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-14 18:10:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-13 18:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-13 19:15:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-15 14:25:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-15 21:41:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-15 07:20:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-15 08:04:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-16 23:06:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-16 23:40:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-17 14:34:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-17 14:42:30", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-04 17:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-04 17:59:00", tz = "Europe/Zurich") | 
  posmo$datetime > as.POSIXct("2023-06-03 21:50:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-03 22:18:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-06-02 18:02:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-06-02 18:43:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-06 21:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-06 22:37:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-09 09:50:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-09 12:05:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-10 08:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-10 09:39:00", tz = "Europe/Zurich") |
  posmo$datetime > as.POSIXct("2023-04-10 18:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-10 21:10:00", tz = "Europe/Zurich")
  
#labelize data
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_cars ~ 'car'))

# Visualisation of all labelized tmode datasets
posmo_cars <- posmo %>% filter(tmode_manual == "car") %>% filter(region == "Waadt") #Waadt #just for visualization, 

#plot trajectories
ggplot(posmo_cars, aes(E,N, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## test
```{r}
# to see segments
  
#car
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-17 07:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-17 07:57:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-17 10:03:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-17 10:07:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-17 16:04:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-17 16:21:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-17 18:26:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-17 18:44:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-18 08:32:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-18 08:37:00", tz = "Europe/Zurich"))
posmo_test

#tram
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-18 08:39:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-18 08:47:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-18 08:47:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-18 08:55:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-18 17:10:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-18 17:15:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-18 17:17:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-18 17:40:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-18 17:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-18 17:48:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-18 17:55:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-18 18:05:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-18 19:35:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-18 19:51:30", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 08:32:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 08:39:00", tz = "Europe/Zurich"))
posmo_test

#tram
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 08:41:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 08:46:30", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 08:47:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 08:55:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 18:27:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 18:33:30", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 18:34:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 18:43:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 19:34:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 19:49:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 19:49:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 19:55:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 21:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 21:59:30", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-19 22:01:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-19 22:15:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 06:44:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 06:48:00", tz = "Europe/Zurich"))
posmo_test

#train
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 06:51:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 06:54:00", tz = "Europe/Zurich"))
posmo_test

#train
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 07:12:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 07:31:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 07:35:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 07:40:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 07:40:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 07:48:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 12:02:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 12:47:00", tz = "Europe/Zurich"))
posmo_test

#train
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 12:48:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 13:13:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 13:13:30", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 13:30:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 13:50:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 14:00:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 15:19:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 15:36:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 16:51:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 16:57:00", tz = "Europe/Zurich"))
posmo_test

#train
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 18:24:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 19:52:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 19:53:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 19:59:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 19:59:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 20:15:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-21 22:48:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-21 22:58:00", tz = "Europe/Zurich"))
posmo_test

#car
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-22 03:45:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-22 04:05:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-22 14:20:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-22 14:25:00", tz = "Europe/Zurich"))
posmo_test

#train
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-22 14:30:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-22 14:58:00", tz = "Europe/Zurich"))
posmo_test

#tram
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-22 15:21:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-22 15:40:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-22 18:00:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-22 18:16:00", tz = "Europe/Zurich"))
posmo_test

#bus
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-22 18:16:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-22 18:45:00", tz = "Europe/Zurich"))
posmo_test

#walk
posmo_test <- posmo |>
  filter(posmo$datetime > as.POSIXct("2023-04-22 18:46:00", tz = "Europe/Zurich") & posmo$datetime < as.POSIXct("2023-04-22 19:00:00", tz = "Europe/Zurich"))
posmo_test

posmo_wgs <- posmo_test %>% st_transform(crs = 4326)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(data = posmo_wgs,
                   #opacity = 0.3, 
                   radius = 0.2,
                   popup = posmo_wgs$datetime,
                   color = 'blue') %>%
  addLegend(position = 'topright', 
            colors = c('blue'), 
            labels = c('points'),
            title = 'Downloaded points from posmo')

# Plot the leaflet object m
m

# #plot trajectories
# ggplot(posmo_test, aes(E,N, color = datetime)) +
#   geom_point() +
#   geom_path() +
#   coord_equal() #so equally space North and East
```

## Final Labelizing

```{r}
#final labelizing
posmo <- posmo %>%
  mutate(tmode_manual = case_when(condition_walks ~ 'walk',
                                  condition_boats ~ 'boat',
                                  condition_trains ~ 'train',
                                  condition_bikes ~ 'bike',
                                  condition_scooters ~ 'kick-scooter',
                                  condition_cars ~ 'car',
                                  condition_bus ~ 'bus',
                                  condition_trams ~ 'tram',
                                  condition_t_bars ~ 't_bar',
                                  condition_cable_cars ~ 'cable_car',
                                  condition_ski_lifts ~ 'ski_lift',
                                  condition_skis ~ 'ski'
                                  ))


posmo %>% group_by(tmode_manual) %>% count() %>% View()


posmo_filter <- posmo %>%
  #filter(as.Date(datetime) == "2023-04-25") %>%
  #filter(is.na(tmode_manual))
  #filter(as.Date(datetime) == "2023-04-23") 
  filter(tmode_manual == "walk") #%>% filter(as.Date(datetime) == "2023-04-15") 
#filter(!is.na(tmode_manual)) #%>% filter(region == "Zurich") 


posmo_wgs <- posmo_filter %>% st_transform(crs = 4326)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addCircleMarkers(data = posmo_wgs,
                   #opacity = 0.3, 
                   radius = 0.2,
                   popup = posmo_wgs$datetime,
                   color = 'blue') %>%
  addLegend(position = 'topright', 
            colors = c('blue'), 
            labels = c('points'),
            title = 'Downloaded points from posmo')

# Plot the leaflet object m
m


```


# Qualify our labelized dataset 
Deriving speed: Ex2 - task5 (rolling window)  

